Using simulation to optimize the sampling design of fisheries-independent trawl surveys

### 2018-07-17 analysis/paper/sim_survey_paper.Rmd
Add fancharts to paper (replace "worm plots")

@@

```{r est-fit, cache = TRUE, fig.height = 8, fig.width = 7, fig.cap = "- Stratified estimates of abundance at age across years from 10 simulations of a survey with a set density of 0.002 sets / km^2^ and sampling caps of 500 lengths per set and 10 ages per length group. The heavy dark line indicates the true tred in the population available to the survey."}

which_survey <- 98

d <- sim$age_strat_error
d <- d[d$survey == which_survey & d$sim %in% 1:10, ]
d$age[d$age > 8] <- 8 # aggregate beyond 10
d$age <- as.character(d$age)
d$age[d$age == "8"] <- "8+"
d$age <- factor(d$age, levels = c(as.character(1:7), "8+"))
d <- d[, list(I_hat = sum(I_hat), I = sum(I)), by = c("year", "age", "sim")]

p <- ggplot(data = d) + 
  geom_line(aes(x = year, y = I_hat, color = sim, group = sim), size = 0.1) +
  geom_line(aes(x = year, y = I), color = viridis::viridis(1), size = 0.7) +
  scale_color_viridis(begin = 0.5, guide = FALSE) +
  facet_wrap(~age, scales = "free_y", dir = "v", nrow = 4) +
  xlab("Year") + ylab("Abundance index") +
  theme_minimal() +
  scale_y_continuous(labels = format_si()) +
  scale_x_continuous(expand = c(0.01, 0)) +
  theme(legend.justification = c(0, 1),
        plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm"))

cowplot::save_plot(plot = p, filename = "analysis/paper/figures/true_vs_est.pdf",
                   base_height = 8, base_width = 7)
p

@@

@@

@@

p <- wireframe(RMSE ~  lengths_cap * ages_cap | set_den, data = d,
          as.table = TRUE, drape = TRUE, col = "white", lwd = 0.5,
          strip = strip.custom(factor.levels = labs),
          scales = list(arrows = FALSE, cex = 0.7, z = zlabs),
          col.regions = viridis::viridis(100),
          screen = list(z = 230, x = 300), 
          xlab = list(expression(M[lengths]), rot = 25), 
          ylab = list(expression(M[ages]), rot = -15), 
          zlab = list("RMSE", rot = 95),
          colorkey = list(labels = zlabs, height = 0.6,
                          axis.line = list(col = "darkgrey")),
          par.settings = trellis_theme)
p <- update(p, layout = c(3, 1))
p

dev.off()

p

@@

```{r age-fig, fig.width = 7, fig.height = 6}

## quantiles for fan plot
quants <- seq(90, 10, by = -10)
ints <- lapply(quants, function(q) {
  d[, list(prob = paste0(q, "%"),
           lower = quantile(I_hat, prob = (1 - q / 100) / 2),
           upper = quantile(I_hat, prob = 1 - (1 - q / 100) / 2)), 
    by = c("year", "age", "set_den", "lengths_cap")]
})
ints <- rbindlist(ints)
ints$prob <- factor(ints$prob, levels = paste0(quants, "%"))

true <- d[d$sim == 1, ]

cols <- viridis::viridis(3)
ggplot(data = ints) + 
  geom_ribbon(aes(x = age, ymin = lower, ymax = upper, fill = prob, alpha = prob)) +
  geom_line(aes(x = age, y = I), data = d, color = cols[1], size = 1) +
  scale_fill_viridis(name = "Pr", discrete = TRUE, begin = 0.1) +
  scale_alpha_discrete(name = "Pr") + 
  facet_grid(set_den ~ lengths_cap) +
  xlab("Age") + ylab("Abundance index") +
  theme_minimal() +
  scale_y_continuous(labels = format_si()) +
  scale_x_continuous(expand = c(0.01, 0), 
                     labels = c("2", "4", "6", "8+"), 
                     breaks = seq(2, 8 , 2)) +
  guides(color = guide_legend(keywidth = 0.4, keyheight = 0.25, 
                              default.unit = "cm", ncol = 1)) +
  theme(legend.justification = c(0, 1),
        plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm"))

```{r year-fig, fig.width = 7, fig.height = 6}

## quantiles for fan plot
quants <- seq(90, 10, by = -10)
ints <- lapply(quants, function(q) {
  d[, list(prob = paste0(q, "%"),
           lower = quantile(I_hat, prob = (1 - q / 100) / 2),
           upper = quantile(I_hat, prob = 1 - (1 - q / 100) / 2)), 
    by = c("year", "age", "set_den", "lengths_cap")]
})
ints <- rbindlist(ints)
ints$prob <- factor(ints$prob, levels = paste0(quants, "%"))

true <- d[d$sim == 1, ]

cols <- viridis::viridis(3)
ggplot(data = ints) + 
  geom_ribbon(aes(x = year, ymin = lower, ymax = upper, fill = prob, alpha = prob)) +
  geom_line(aes(x = year, y = I), data = d, color = cols[1], size = 1) +
  scale_fill_viridis(name = "Pr", discrete = TRUE, begin = 0.1) +
  scale_alpha_discrete(name = "Pr") + 
  facet_grid(set_den ~ lengths_cap) +
  xlab("Year") + ylab("Abundance index") +
  theme_minimal() +
  scale_y_continuous(labels = format_si()) +
  scale_x_continuous(expand = c(0.01, 0)) +
  guides(color = guide_legend(keywidth = 0.4, keyheight = 0.25, 
                              default.unit = "cm", ncol = 1)) +
  theme(legend.justification = c(0, 1),
        plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm"))

```{r tot-fig, fig.width = 7, fig.height = 3}

## quantiles for fan plot
quants <- seq(90, 10, by = -10)
ints <- lapply(quants, function(q) {
  d[, list(prob = paste0(q, "%"),
           lower = quantile(I_hat, prob = (1 - q / 100) / 2),
           upper = quantile(I_hat, prob = 1 - (1 - q / 100) / 2)), 
    by = c("year", "set_den")]
})
ints <- rbindlist(ints)
ints$prob <- factor(ints$prob, levels = paste0(quants, "%"))

true <- d[d$sim == 1, ]

cols <- viridis::viridis(3)
ggplot(data = ints) + 
  geom_ribbon(aes(x = year, ymin = lower, ymax = upper, fill = prob, alpha = prob)) +
  geom_line(aes(x = year, y = I), data = d, color = cols[1], size = 1) +
  scale_fill_viridis(name = "Pr", discrete = TRUE, begin = 0.1) +
  scale_alpha_discrete(name = "Pr") + 
  facet_wrap(~ set_den, nrow = 1) +
  xlab("Year") + ylab("Abundance index") +
  theme_minimal() +
  scale_y_continuous(labels = format_si()) +
  scale_x_continuous(expand = c(0.01, 0)) +
  guides(color = guide_legend(keywidth = 0.4, keyheight = 0.25, 
                              default.unit = "cm", ncol = 1)) +
  theme(legend.justification = c(0, 1),
        plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm"))

## Figure to-dos

- Fix facet labels (make them look like the wireframe plot)
- Add RMSE label to colorkey in wireframe plot
- Change vis_sim estimates plot to fan plots like used here

>>>>

## little helper function for the fan plot
fan_plot <- function(d, x = "Age", total_strat = FALSE) {

  if (total_strat) {
    ints_by = c("year", "set_den")
  } else {
    ints_by = c("year", "age", "set_den", "lengths_cap")
  }

  ## quantiles for fan plot
  quants <- seq(90, 10, by = -10)
  ints <- lapply(quants, function(q) {
    d[, list(prob = paste0(q, "%"),
             lower = quantile(I_hat, prob = (1 - q / 100) / 2),
             upper = quantile(I_hat, prob = 1 - (1 - q / 100) / 2)), 
      by = ints_by]
  })
  ints <- rbindlist(ints)
  ints$prob <- factor(ints$prob, levels = paste0(quants, "%"))

  true <- d[d$sim %in% 1:3, ]

  if (x == "Age") {
    p <- ggplot(data = ints, aes(x = age))
    xlabs <- c("2", "4", "6", "8+")
    xbreaks <- seq(2, 8 , 2)
  }
  if (x == "Year") {
    p <- ggplot(data = ints, aes(x = year))
    xlabs <- waiver()
    xbreaks <- waiver()
  }

  p <-  p + 
    geom_ribbon(aes(ymin = lower, ymax = upper, fill = prob, alpha = prob)) +
    geom_line(aes(y = I), data = true, color = viridis::viridis(1), size = 0.7) +
    scale_fill_viridis(name = "Pr", discrete = TRUE, begin = 0.1) +
    scale_alpha_discrete(name = "Pr") + 
    xlab(x) + ylab("Abundance index") +
    scale_y_continuous(labels = format_si()) +
    scale_x_continuous(expand = c(0.01, 0), 
                       labels = xlabs, 
                       breaks = xbreaks) +
    guides(color = guide_legend(keywidth = 0.4, keyheight = 0.25, 
                                default.unit = "cm", ncol = 1))

  if (total_strat) {
    p <- p + facet_wrap(~ set_den, nrow = 1,
                        labeller = label_bquote(cols = D[sets]==.(set_den)))
  } else {
    p <- p + facet_grid(lengths_cap ~ set_den, 
                        labeller = label_bquote(cols = D[sets]==.(set_den),
                                                rows = M[lengths]==.(lengths_cap)))
  }

  p + theme_bw() +
    theme(legend.justification = c(0, 1),
          plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm"),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          strip.background = element_blank(),
          panel.border = element_rect(fill = NA, color = "darkgrey", size = 0.4),
          axis.ticks = element_line(color = "darkgrey", size = 0.4))

}

@@

\pagebreak

```{r tot-fig, cache = TRUE, fig.width = 8, fig.height = 3, fig.cap = "- Fanchart of stratified estimates of the trend in total abundance from surveys with different set densities, $D_{sets}$. The heavy dark line indicates the true tred in the total population available to the survey."}

p <- fan_plot(d, x = "Year", total_strat = TRUE)
cowplot::save_plot(plot = p, filename = "analysis/paper/figures/total_strat_true_vs_est.pdf",
                   base_height = 3, base_width = 8)
p

@@

\pagebreak

```{r age-fig, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fanchart of stratified estimates of abundance at age from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$. Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios. The heavy dark line indicates the true population available to the survey."}

p <- fan_plot(d, x = "Age")
cowplot::save_plot(plot = p, filename = "analysis/paper/figures/true_vs_est_by_age.pdf",
                   base_height = 6, base_width = 8)
p

\pagebreak

```{r year-fig, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fanchart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$. Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios. The heavy dark line indicates the true tred in the population available to the survey."}

p <- fan_plot(d, x = "Year")
cowplot::save_plot(plot = p, filename = "analysis/paper/figures/true_vs_est_by_year.pdf",
                   base_height = 6, base_width = 8)
p

\pagebreak

@@

# wireframe_legend <- packGrob(
#   draw.colorkey(key = list(col = viridis::viridis, at = do.breaks(range(z),100))), 
#   textGrob("RMSE", x = 0, y = 0.5, just = c("left", "centre")), 
#   height = unit(2, "lines"), side = "top", dynamic = TRUE)

@@

p <- wireframe(RMSE ~  lengths_cap * ages_cap | set_den, data = d,
          as.table = TRUE, drape = TRUE, col = "white", lwd = 0.5,
          strip = strip.custom(factor.levels = labs),
          scales = list(arrows = FALSE, cex = 0.7, z = zlabs),
          col.regions = viridis::viridis(100),
          screen = list(z = 230, x = 300), 
          xlab = list(expression(M[lengths]), rot = 25), 
          ylab = list(expression(M[ages]), rot = -15), 
          zlab = list("RMSE", rot = 95),
          colorkey = list(labels = zlabs, height = 0.6,
                          axis.line = list(col = "darkgrey")),
          par.settings = trellis_theme)
p <- update(p, layout = c(3, 1))
p
grid::grid.text("RMSE", x = unit(0.952, "npc"), y = unit(0.77, "npc"))

invisible(dev.off())

p
grid::grid.text("RMSE", x = unit(0.952, "npc"), y = unit(0.77, "npc"))

@@

\pagebreak
